---
title: "Explore year-week-height averages"
date: "`r Sys.Date()`"
author: "Danielle Dempsey"
format: 
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
echo: false
editor_options: 
  chunk_output_type: console
---

```{r}
#| include: false

library(data.table)
library(dplyr)
library(ggplot2)
library(here)
library(htmltools)
library(lubridate)
library(plotly)
library(RColorBrewer)
library(viridis)

source(here("contributor_folders/danielle/functions/helpers.R"))

theme_set(theme_light())

# ensures the Plotly JS dependencies are loaded before being called in the loop
htmltools::tagList(plot_ly(x = rnorm(10), type = "histogram"))


path <- path <- "/home/jovyan/shared-public/ohw26/model_downsampling/"

dat <- fread(
  paste0(path, "cmar_weekly_average_temperature.csv"), data.table = FALSE
) %>% 
  mutate(week_date = make_date_from_iso_year_week(iso_year, iso_week)) 

stations <- unique(dat$station)

timeline_dat <- fread( paste0(path, "temperature_observations_timeline.csv"), data.table = FALSE)

```

This file visualizes the average and standard deviation of temperature calculated for each year, week, station, and sensor height.

This is mainly to get a qualitative sense of how the average and standard deviation change over time and space, and to identify any obvious errors or outliers.

Some figures are interactive, but some are not (depending on the size and complexity).

A separate file will assess the inter-annual variability of the weekly averages to determine if the average weekly temperature can be used as a representative time series for comparison to the FVCOM model.

# Timeline

None of the stations have observed data for the period of the FCOM model run.

```{r}
#| warning: false

pal_foo <- colorRampPalette(brewer.pal(8, "Dark2"))
pal <- pal_foo(length(stations) + 1)

p <- ggplot(timeline_dat) +
  geom_segment(
    aes(
      x = start_date, xend = end_date, 
      y = station, yend = station, col = station,
      text = paste(
        "station: ", station, "<br>",
        "start_date: ", start_date, "<br>",
        "end_date: ", end_date, "<br>",
        "n weeks: ", ts_length_weeks
      )
    ), linewidth = 4
  ) +
  scale_y_discrete(limits = rev) +
  scale_colour_manual(values = pal, drop = FALSE) +
  theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.border = element_rect(colour = "gray20"),
    strip.background = element_rect(colour = "gray20", fill = "gray40"),
    panel.spacing.x = unit(-1, "lines")
  )

ggplotly(p, tooltip = "text")
```


# Average Weekly Temperature

The average temperature was calculated for each year, ISO week, station, and sensor height.

```{=html}
<style>
.panel-tabset .tab-content {
  min-height: 450px;
}
</style>
```

## Temperature Time Series

::: {.panel-tabset}

```{r}
#| results: asis
#| warning: false

for(i in seq_along(stations)) {

  station_i <- stations[i]

  cat('\n###', station_i, '\n')
  
  dat_i <- dat %>% 
    filter(station == station_i) %>% 
    # so a discrete colour scale is used for sensor height
    mutate(
      sensor_height_above_seafloor_m =  convert_to_ordered_factor(sensor_height_above_seafloor_m)
    )
  
  pal <- viridis(
    length(unique(dat_i$sensor_height_above_seafloor_m)), option = "D", direction = -1
  )
  
  p <- ggplot(
    dat_i, aes(week_date, 
               mean_temperature_degree_c, color = sensor_height_above_seafloor_m)) +
    geom_point() +
    scale_colour_manual("sensor height (m)", values = pal) +
    scale_y_continuous("temperature_degree_c", limits = c(-2, 20)) +
    scale_x_date(limits = c(as_date("2016-01-01"), as_date("2025-12-31")))
  
   print(htmltools::tagList(ggplotly(p)))
}

```
:::

## Average Temperature +/- Standard Deviation

- Black points are the average temperature.
- Ribbons show average temperature +/- standard deviation.
- Note differences in x-axis, y-axis, and colour scales.

::: {.panel-tabset}

```{r}
#| results: asis
#| warning: false

for(i in seq_along(stations)) {
  
  station_i <- stations[i]
  
  cat('\n###', station_i, '\n')
  
  dat_i <- dat %>% 
    filter(station == station_i) %>% 
    mutate(
      sensor_height_above_seafloor_m =  convert_to_ordered_factor(sensor_height_above_seafloor_m)
    )
  
  pal <- viridis(
    length(unique(dat_i$sensor_height_above_seafloor_m)), option = "D", direction = -1
  )
  
  p <- ggplot(dat_i) +
    geom_ribbon(
      aes(x = week_date,
          ymin = mean_temperature_degree_c - sd_temperature_degree_c,
          ymax = mean_temperature_degree_c + sd_temperature_degree_c,
          color = sensor_height_above_seafloor_m,
          fill = sensor_height_above_seafloor_m
      ), alpha = 0.5
    ) +
    geom_point(aes(week_date, mean_temperature_degree_c), size = 0.25) +
    scale_colour_manual("sensor height (m)", values = pal) +
    scale_fill_manual("sensor height (m)", values = pal) +
    facet_wrap(~sensor_height_above_seafloor_m, ncol = 1) 
  
  n_height <- length(unique(dat_i$sensor_height_above_seafloor_m))
  
  if(n_height == 1) {
    h <- 2.5
  } else h <- n_height * 2.25
  
  subchunkify(p, fig_width = 8, fig_height = h)
}

```
::: 


# Standard Deviation

## Time Series

How much variability is there within the observations used to calculate each weekly average value?

::: {.panel-tabset}
```{r}
#| results: asis
#| warning: false

for(i in seq_along(stations)) {

  station_i <- stations[i]

  cat('\n###', station_i, '\n')
  
  dat_i <- dat %>% 
    filter(station == station_i) %>% 
    mutate(sensor_height_above_seafloor_m = convert_to_ordered_factor(sensor_height_above_seafloor_m))
  
  pal <- viridis(
    length(unique(dat_i$sensor_height_above_seafloor_m)), option = "D", direction = -1
  )
  
  p <- ggplot(
    dat_i, aes(week_date, 
               sd_temperature_degree_c, color = sensor_height_above_seafloor_m)) +
    geom_point() +
    scale_colour_manual("sensor height (m)", values = pal) +
    scale_y_continuous("temperature_degree_c", limits = c(0, 2)) 
 
   print(htmltools::tagList(ggplotly(p)))
}
```
:::

## Boxplots by Year & Sensor Height

How does the standard deviation change over the observed years for each year-week-station-height group?

::: {.panel-tabset}
```{r}
#| results: asis
#| warning: false

for(i in seq_along(stations)) {

  station_i <- stations[i]

  cat('\n###', station_i, '\n')
  
  dat_i <- dat %>% 
    filter(station == station_i) %>% 
    mutate(sensor_height_above_seafloor_m = convert_to_ordered_factor(sensor_height_above_seafloor_m))
  
  pal <- viridis(
    length(unique(dat_i$sensor_height_above_seafloor_m)), option = "D", direction = -1
  )
  
  p <- ggplot(
    dat_i, 
    aes(iso_year, sd_temperature_degree_c, fill = sensor_height_above_seafloor_m, 
        group = iso_year)) +
    geom_boxplot(outliers = TRUE) +
    scale_y_continuous(limits = c(0, 2)) +
    scale_x_continuous(breaks = seq(min(dat_i$iso_year), max(dat_i$iso_year))) +
    scale_fill_manual("sensor height (m)", values = pal) + 
    facet_wrap(~sensor_height_above_seafloor_m, scales = "free_x", ncol = 2) +
    theme(legend.position = "none")
  
  n_height <- length(unique(dat_i$sensor_height_above_seafloor_m))
  if(n_height <= 2) {
    h <- 2
  } else h <- n_height / 2 * 2.25
  
  subchunkify(p, fig_width = 8, fig_height = h)
}

```
:::

## Boxplots by Sensor Height

How does the standard deviation change over the sensor height for each year-week-station-height group?

::: {.panel-tabset}
```{r}
#| results: asis
#| warning: false

for(i in seq_along(stations)) {

  station_i <- stations[i]

  cat('\n###', station_i, '\n')
  
  dat_i <- dat %>% 
    filter(station == station_i) %>% 
    mutate(
      sensor_height_above_seafloor_m =  convert_to_ordered_factor(sensor_height_above_seafloor_m)
    )
  
  pal <- viridis(
    length(unique(dat_i$sensor_height_above_seafloor_m)), option = "D", direction = -1
  )
  
  p <- ggplot(
    dat_i, aes(sensor_height_above_seafloor_m, sd_temperature_degree_c, fill = sensor_height_above_seafloor_m)) +
    geom_boxplot(outliers = TRUE) +
    scale_fill_manual("sensor height (m)", values = pal) +
    scale_y_continuous(limits = c(0, 2)) +
    theme(legend.position = "none")
  
  print(htmltools::tagList(ggplotly(p)))
}

```
:::


# Numer of Observations

How many observations were included in each average temperature value?

- Could consider removing the first and last week of each deployment, since fewer observations are recorded. This likely does not impact the average. 
  - Review the average temperature time series to see.
- In 2024, one sensor at Long Island 1 and one sensor at Sandy Cove were each set to record every 1 minute, which is where there are ~10,000 observations each week.

::: {.panel-tabset}
```{r}
#| results: asis
#| warning: false

for(i in seq_along(stations)) {

  station_i <- stations[i]

  cat('\n###', station_i, '\n')
  
  dat_i <- dat %>% 
    filter(station == station_i) %>% 
    mutate(
      sensor_height_above_seafloor_m =  convert_to_ordered_factor(sensor_height_above_seafloor_m)
    )
  
  pal <- viridis(
    length(unique(dat_i$sensor_height_above_seafloor_m)), option = "D", direction = -1
  )
  
  p <- ggplot(
    dat_i, aes(week_date, n, col = sensor_height_above_seafloor_m)) +
    geom_point(size = 0.5) +
    scale_colour_manual("sensor height (m)", values = pal) 
  
  print(htmltools::tagList(ggplotly(p)))
}

```
:::

# Potential code improvements

  - `map` over stations instead of loop to reduce repetition.
  - divide figures by station instead of figure type.

