---
title: "2.2: Compare Weekly Temperatures between Years"
date: "`r Sys.Date()`"
author: "Danielle Dempsey"
format: 
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
echo: false
editor_options: 
  chunk_output_type: console
---

```{r}
#| include: false

library(data.table)
library(dplyr)
library(DT)
library(ggplot2)
library(here)
library(htmltools)
library(leaflet)
library(lubridate)
library(plotly)
library(RColorBrewer)
library(viridis)

source(here("contributor_folders/danielle/functions/helpers.R"))

theme_set(theme_light())

# ensures the Plotly JS dependencies are loaded before being called in the loop
htmltools::tagList(plot_ly(x = rnorm(10), type = "histogram"))


path <- path <- "/home/jovyan/shared-public/ohw26/model_downsampling/"

dat <- fread(
  paste0(path, "cmar_average_temperature_year_week.csv"), data.table = FALSE
)

dat_summary <- dat %>% 
  summarise(
    wk_mean_temperature_degree_c = mean(mean_temperature_degree_c),
    wk_sd_temperature_degree_c = sd(mean_temperature_degree_c),
    n = n(),
    .by = c(iso_week, station, sensor_height_above_seafloor_m)
  )

# save this file for use in other scripts. "rep" stands for "representative"
# fwrite(dat_summary, paste0(path, "cmar_rep_weekly_average_temperatures.csv"))

stations <- unique(dat$station)
```


```{=html}
<style>
.panel-tabset .tab-content {
  min-height: 450px;
}
</style>
```


# Weekly Average Temperature

- By year, station, and height.

::: {.panel-tabset}

```{r}
#| results: asis
#| warning: false

for(i in seq_along(stations)) {

  station_i <- stations[i]

  cat('\n###', station_i, '\n')
  
  dat_i <- dat %>% 
    filter(station == station_i) %>%
    mutate(
      sensor_height_above_seafloor_m = convert_to_ordered_factor(sensor_height_above_seafloor_m),
      year = factor(iso_year)
    )
  
  pal <- brewer.pal(length(unique(dat_i$year)), "Dark2")
  
  p <- ggplot(dat_i, aes(iso_week, mean_temperature_degree_c, col = year)) +
    geom_point() + 
    scale_color_manual(values = pal) +
    scale_x_continuous(limits = c(1, 53)) +
    facet_wrap(~sensor_height_above_seafloor_m, ncol = 2)
  
  n_height <- length(unique(dat_i$sensor_height_above_seafloor_m))
  n_row <- ceiling(n_height / 2)
  
  if(n_row == 1) {
    h <- 2
  } else h <- n_row * 2
  
  subchunkify(p, fig_width = 8, fig_height = h)
}

```
:::

# Representative Average

- This is the average of the weekly average temperatues over all observed years.
- Ribbons show average temperature +/- standard deviation.
- No ribbon around a point indicates that there is only one year of data for that ISO week.
- Gaps indicate that data was not recorded in that ISO week for any years.

::: {.panel-tabset}

```{r}
#| results: asis
#| warning: false
#| message: false

for(i in seq_along(stations)) {
  
  station_i <- stations[i]
  
  cat('\n###', station_i, '\n')
  
  dat_i <- dat_summary %>% 
    filter(station == station_i) %>%
    mutate(
      sensor_height_above_seafloor_m = convert_to_ordered_factor(sensor_height_above_seafloor_m)
    )
  
  pal <- viridis(
    length(unique(dat_i$sensor_height_above_seafloor_m)), option = "D", direction = -1
  )
  
  p <- ggplot(dat_i) +
      geom_ribbon(
      aes(x = iso_week,
          ymin = wk_mean_temperature_degree_c - wk_sd_temperature_degree_c,
          ymax = wk_mean_temperature_degree_c + wk_sd_temperature_degree_c,
          color = sensor_height_above_seafloor_m,
          fill = sensor_height_above_seafloor_m
      ), alpha = 0.4, na.rm = TRUE
    ) +
    geom_point(aes(iso_week, 
                   wk_mean_temperature_degree_c, 
                   col = sensor_height_above_seafloor_m)) +
    scale_colour_manual("sensor height (m)", values = pal) +
    scale_fill_manual("sensor height (m)", values = pal) +
    scale_x_continuous(limits = c(1, 53)) +
    facet_wrap(~sensor_height_above_seafloor_m, ncol = 2) 
  
  n_height <- length(unique(dat_i$sensor_height_above_seafloor_m))
  n_row <- ceiling(n_height / 2)
  
  if(n_row == 1) {
    h <- 2
  } else h <- n_row * 2
  
  subchunkify(p, fig_width = 8, fig_height = h)
}

```
:::

# Standard Deviation 

- Only includes stations with more than one year of data for at least one week. (The standard deviation is `NA` if there is only one year of data.)

## By Station

::: panel-tabset

### Boxplot
```{r}
dat_sd <- dat_summary %>% 
  filter(!is.na(wk_sd_temperature_degree_c))

p <- ggplot(dat_sd, aes(station, wk_sd_temperature_degree_c, fill = station)) + 
  geom_boxplot(outliers = TRUE) +
  theme(legend.position = "none")

ggplotly(p)
```

### Distribution
```{r}

p <- ggplot(dat_sd, aes(wk_sd_temperature_degree_c, fill = station)) +
   geom_histogram(binwidth = 0.2, color = "grey80", linewidth = 0.25)

ggplotly(p)

```

### Table

```{r}
dat_sd %>% 
  summarise(
    min_sd = round(min(wk_sd_temperature_degree_c), digits = 3),
    med_sd = round(median(wk_sd_temperature_degree_c), digits = 3),
    mean_sd = round(mean(wk_sd_temperature_degree_c), digits = 3),
    max_sd = round(max(wk_sd_temperature_degree_c), digits = 3),
    .by = station
  ) %>% 
  datatable(
    rownames = FALSE, 
    options = list(
      dom = 'Bft',
      paging = FALSE,
      searching = TRUE,
      pageLength = 500,
      scrollY = "500px",
      scrollX = "500px",
      columnDefs = list(list(className = 'dt-center', targets = "_all")),
      buttons = 'copy'
    ))

```

:::


## Time Series

::: {.panel-tabset}
```{r}
#| results: asis
#| warning: false

for(i in seq_along(stations)) {
  
  station_i <- stations[i]
  
  dat_i <- dat_summary %>% 
    filter(station == station_i) %>% 
    mutate(sensor_height_above_seafloor_m = convert_to_ordered_factor(sensor_height_above_seafloor_m))
  
  if(sum(is.na(dat_i$wk_sd_temperature_degree_c)) != nrow(dat_i)) {
    cat('\n###', station_i, '\n')
    
    pal <- viridis(
      length(unique(dat_i$sensor_height_above_seafloor_m)), option = "D", direction = -1
    )
    
    p <- ggplot(
      dat_i, aes(iso_week, 
                 wk_sd_temperature_degree_c, color = sensor_height_above_seafloor_m)) +
      geom_point() +
      scale_colour_manual("sensor height (m)", values = pal) +
      scale_y_continuous("temperature_degree_c", limits = c(0, 2)) 
    
    print(htmltools::tagList(ggplotly(p)))
  }
}

```
:::


## Boxplots by Sensor Height

- Only shows stations with more than one year of data for at least one week. (The standard deviation is `NA` if there is only one year of data.)

::: {.panel-tabset}
```{r}
#| results: asis
#| warning: false

for(i in seq_along(stations)) {

  station_i <- stations[i]

   dat_i <- dat_summary %>% 
    filter(station == station_i) %>% 
    mutate(
      sensor_height_above_seafloor_m =  convert_to_ordered_factor(sensor_height_above_seafloor_m)
    )
  
   if(sum(is.na(dat_i$wk_sd_temperature_degree_c)) != nrow(dat_i)) {
     cat('\n###', station_i, '\n')
     
     pal <- viridis(
       length(unique(dat_i$sensor_height_above_seafloor_m)), option = "D", direction = -1
     )
     
     p <- ggplot(
       dat_i, aes(
         sensor_height_above_seafloor_m, wk_sd_temperature_degree_c, 
         fill = sensor_height_above_seafloor_m)
     ) +
       geom_boxplot(outliers = TRUE) +
       scale_fill_manual("sensor height (m)", values = pal) +
       scale_y_continuous(limits = c(0, 2)) +
       theme(legend.position = "none")
     
     print(htmltools::tagList(ggplotly(p)))
   }
}

```
:::

# Map

- Centreville, Long Beach, Long Island 2, and Sandy Cove are the only stations with at least one **full** year of data. 


```{r}

st_dat <- dat %>%
  distinct(station, station_elevation_m, latitude, longitude) %>% 
  left_join(
    dat_summary %>% 
      summarise(min_n_years =  min(n), max_n_years = max(n), .by = station),
    by = "station"
  ) %>% 
  mutate(
    popup = paste(
      "station: ", station, "</br>",
      "elevation: ", station_elevation_m, "</br>",
      "min_n_years: ", min_n_years, "</br>",
      "max_n_years: ", max_n_years)
  )


st_pal <- colorFactor(
  palette = brewer.pal(length(unique(st_dat$max_n_years)), "Dark2"),
  domain = unique(st_dat$max_n_years)
)
st_map <- split(st_dat, st_dat$max_n_years)


leaflet() %>%
  addProviderTiles("Esri.WorldTopoMap") %>% 
  add_map_layers(
    map_dat = st_map,
    popup = ~popup,
    size = 5,
    map_pal = ~st_pal(max_n_years)
  ) %>% 
  addScaleBar(
    position = "topright",
    options = scaleBarOptions(imperial = FALSE)
  ) %>%
  addLayersControl(
    overlayGroups = unique(st_dat$max_n_years),
    options = layersControlOptions(collapsed = FALSE),
    position = "bottomright"
  ) %>%
   addLegend(
      "bottomright",
      pal = st_pal,
      values = unique(st_dat$max_n_years),
      title = "Max # Years",
      opacity = 0.75,
      position = "bottomright"
    ) %>%
  htmlwidgets::onRender(
    "function() {$('.leaflet-control-layers-overlays').
    prepend('<label style=\"text-align:center\">Max # Years</label>');}"
  )
```



