---
title: "2.3: Explore `week` Averages"
date: "`r Sys.Date()`"
author: "Danielle Dempsey"
format: 
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
echo: false
editor_options: 
  chunk_output_type: console
---

```{r}
#| include: false

library(data.table)
library(dplyr)
library(DT)
library(ggplot2)
library(here)
library(htmltools)
library(lubridate)
library(plotly)
library(RColorBrewer)
library(viridis)

source(here("contributor_folders/danielle/functions/helpers.R"))

theme_set(theme_light())

# ensures the Plotly JS dependencies are loaded before being called in the loop
htmltools::tagList(plot_ly(x = rnorm(10), type = "histogram"))


path <- "/home/jovyan/shared-public/ohw26/model_downsampling/"

dat <- fread(
  paste0(path, "cmar_average_temperature_week.csv"), data.table = FALSE
) 
stations <- unique(dat$station)

timeline_dat <- fread( paste0(path, "temperature_observations_timeline.csv"), data.table = FALSE)

```

**This is for weekly averages calculated directly from the raw data. Year is not considered.**

  - Advantage: we can calculate standard deviation for stations with only one year of data.
  - Disadvantage: the standard deviation may be more related to intra-week variability (e.g., day-night) than inter-annual variability. We are interested in inter-annual variability of the weekly averages to assess the acceptable ranges of outputs from the model.

This file visualizes the average and standard deviation of temperature calculated for each week, station, and sensor height.

This is mainly to get a qualitative sense of how the average and standard deviation change over time and space, and to identify any obvious errors or outliers.

Some figures are interactive, but some are not (depending on the size and complexity).


# Timeline

None of the stations have observed data for the period of the FCOM model run.

```{r}
#| warning: false

pal_foo <- colorRampPalette(brewer.pal(8, "Dark2"))
pal <- pal_foo(length(stations) + 1)

p <- ggplot(timeline_dat) +
  geom_segment(
    aes(
      x = start_date, xend = end_date, 
      y = station, yend = station, col = station,
      text = paste(
        "station: ", station, "<br>",
        "start_date: ", start_date, "<br>",
        "end_date: ", end_date, "<br>",
        "n weeks: ", ts_length_weeks
      )
    ), linewidth = 4
  ) +
  scale_y_discrete(limits = rev) +
  scale_colour_manual(values = pal, drop = FALSE) +
  theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.border = element_rect(colour = "gray20"),
    strip.background = element_rect(colour = "gray20", fill = "gray40"),
    panel.spacing.x = unit(-1, "lines")
  )

ggplotly(p, tooltip = "text")
```


# Average Weekly Temperature

- The average temperature was calculated for each ISO week, station, and sensor height.
- Ribbon shows +/- standard deviation.

```{=html}
<style>
.panel-tabset .tab-content {
  min-height: 450px;
}
</style>
```

::: {.panel-tabset}

```{r}
#| results: asis
#| warning: false

for(i in seq_along(stations)) {
  
  station_i <- stations[i]
  
  cat('\n###', station_i, '\n')
  
  dat_i <- dat %>% 
    filter(station == station_i) %>% 
    mutate(sensor_height_above_seafloor_m = convert_to_ordered_factor(sensor_height_above_seafloor_m))
  
  pal <- viridis(
    length(unique(dat_i$sensor_height_above_seafloor_m)), option = "D", direction = -1
  )
  
  p <- ggplot(
    dat_i, 
    aes(iso_week, mean_temperature_degree_c, col = sensor_height_above_seafloor_m)
  ) +
    geom_ribbon(
      aes(ymin = mean_temperature_degree_c - sd_temperature_degree_c,
          ymax = mean_temperature_degree_c + sd_temperature_degree_c,
          fill = sensor_height_above_seafloor_m
      ), alpha = 0.4
    ) +
    geom_point() +
    scale_x_continuous(limits = c(1, 53)) +
    scale_colour_manual("sensor height (m)", values = pal) +
    scale_fill_manual("sensor height (m)", values = pal) +
    facet_wrap(~sensor_height_above_seafloor_m, ncol = 2) 
  
  n_height <- length(unique(dat_i$sensor_height_above_seafloor_m))
  n_row <- ceiling(n_height / 2)
  
  if(n_row == 1) {
    h <- 2
  } else h <- n_row * 2
  
  
  subchunkify(p, fig_width = 8, fig_height = h)
}

```
::: 


# Standard Deviation

## By Station

::: panel-tabset

### Boxplot
```{r}
dat_sd <- dat %>% 
  filter(!is.na(sd_temperature_degree_c)) %>% 
  mutate(station = ordered(station, levels = sort(stations, decreasing = TRUE)))

pal <- ggplot_default_colors <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues[1:n], c = 100, l = 65)
}

p <- ggplot(dat_sd, aes(station, sd_temperature_degree_c, fill = station)) + 
  geom_boxplot(outliers = TRUE) +
  coord_flip() +
  scale_fill_manual(values = pal(13)) +
 # scale_y_discrete(limits = rev) +
  theme(legend.position = "none")

ggplotly(p)
```

### Distribution
```{r}

p <- ggplot(dat_sd, aes(sd_temperature_degree_c, fill = station)) +
   geom_histogram(binwidth = 0.2, color = "grey80", linewidth = 0.25)

ggplotly(p)

```

### Table

```{r}
dat_sd %>% 
  summarise(
    min_sd = round(min(sd_temperature_degree_c), digits = 3),
    med_sd = round(median(sd_temperature_degree_c), digits = 3),
    mean_sd = round(mean(sd_temperature_degree_c), digits = 3),
    max_sd = round(max(sd_temperature_degree_c), digits = 3),
    .by = station
  ) %>% 
  datatable(
    rownames = FALSE, 
    options = list(
      dom = 'Bft',
      paging = FALSE,
      searching = TRUE,
      pageLength = 500,
      scrollY = "500px",
      scrollX = "500px",
      columnDefs = list(list(className = 'dt-center', targets = "_all")),
      buttons = 'copy'
    ))

```

:::


## Time Series

::: {.panel-tabset}
```{r}
#| results: asis
#| warning: false

for(i in seq_along(stations)) {

  station_i <- stations[i]

  cat('\n###', station_i, '\n')
  
  dat_i <- dat %>% 
    filter(station == station_i) %>% 
    mutate(sensor_height_above_seafloor_m = convert_to_ordered_factor(sensor_height_above_seafloor_m))
  
  pal <- viridis(
    length(unique(dat_i$sensor_height_above_seafloor_m)), option = "D", direction = -1
  )
  
  p <- ggplot(
    dat_i, aes(iso_week, 
               sd_temperature_degree_c, color = sensor_height_above_seafloor_m)) +
    geom_point() +
    scale_colour_manual("sensor height (m)", values = pal) +
    scale_y_continuous("temperature_degree_c", limits = c(0, 2)) 
 
   print(htmltools::tagList(ggplotly(p)))
}
```
:::


## Boxplots by Sensor Height

::: {.panel-tabset}
```{r}
#| results: asis
#| warning: false

for(i in seq_along(stations)) {

  station_i <- stations[i]

  cat('\n###', station_i, '\n')
  
  dat_i <- dat %>% 
    filter(station == station_i) %>% 
    mutate(
      sensor_height_above_seafloor_m =  convert_to_ordered_factor(sensor_height_above_seafloor_m)
    )
  
  pal <- viridis(
    length(unique(dat_i$sensor_height_above_seafloor_m)), option = "D", direction = -1
  )
  
  p <- ggplot(
    dat_i, aes(sensor_height_above_seafloor_m, sd_temperature_degree_c, fill = sensor_height_above_seafloor_m)) +
    geom_boxplot(outliers = TRUE) +
    scale_fill_manual("sensor height (m)", values = pal) +
    scale_y_continuous(limits = c(0, 2)) +
    theme(legend.position = "none")
  
  print(htmltools::tagList(ggplotly(p)))
}

```
:::


# Numer of Observations

How many observations were included in each average temperature value?

::: {.panel-tabset}
```{r}
#| results: asis
#| warning: false

for(i in seq_along(stations)) {

  station_i <- stations[i]

  cat('\n###', station_i, '\n')
  
  dat_i <- dat %>% 
    filter(station == station_i) %>% 
    mutate(
      sensor_height_above_seafloor_m =  convert_to_ordered_factor(sensor_height_above_seafloor_m)
    )
  
  pal <- viridis(
    length(unique(dat_i$sensor_height_above_seafloor_m)), option = "D", direction = -1
  )
  
  p <- ggplot(
    dat_i, aes(iso_week, n, col = sensor_height_above_seafloor_m)) +
    geom_point(size = 0.5) +
    scale_colour_manual("sensor height (m)", values = pal) 
  
  print(htmltools::tagList(ggplotly(p)))
}

```
:::


